\documentclass[10pt]{article}

\input{../Preamble}

\begin{document}
\title{FESOM2.0: Technical Aspects}
\date{}
\maketitle
\section*{Friday 8/2/2019}
The code is running for much longer than it should. 
\eqnn{
	\Delta t = \frac{1}{96}\text{day} \qquad \text{\# cells} = 38398 \qquad \text{\# vertical layers} = 24
}
Specifically, running for 10 years on 32 CPUs takes $393854s \approx 109hrs$.
\section*{Tuesday 12/2/2019: Test runs}

We run the code for one day on two CPUs with the same mesh and time-step described above:

Total run time = $1439s \approx 24min$. From this we predict that the model, when run for 10 years on 32 CPUs should take $365\times 10/16$ times as long, i.e. $91.25 hrs$. This estimate is of the same order of magnitude as the actual running time, so this makes sense.  \\
\linebreak
Running again for a day but this time on 32 CPUs gives a run time of $106s$. Note that $1439/16 \approx 90$, so this gives approximately the right scaling with CPU number.  

\section*{Wednesday 13/2/2019}

Let us integrate over a day on varying numbers of CPUs with an allocation of $2\text{GB}$ per CPU and the same specifications as above. We display the results below:

\begin{itemize}
\item 2 CPUs:  $1554.704s \approx 26min$. 
\item 4 CPUs: $599.1404s \approx 10min$.
\item 8 CPUs: $423.3595s \approx 7min$.
\item 16 CPUs: $224.6756s \approx 4min$.
\item 32 CPUs: $107.1845s \approx 2min$.
\end{itemize}

Integrating the code over ten days scales as follows:
\begin{itemize}
\item 2 CPUs:  $15624.89\approx 260min$
\item 4 CPUs: $8232.239\approx 137min$.
\item 8 CPUs: $4256.799\approx 71min$.
\item 16 CPUs: $2324.895\approx 39min$.
\item 32 CPUs: $1071.855\approx 18min$.
\end{itemize}

\section{Results from long run}

In the run with the aforementioned parameters over ten physical years we find the following: 

\section{Model Set-Up}

The equations of motion being solved are:\\
\linebreak
We use a wind forcing of a latitude-dependent cosine wave, constant in time and longitude. Something like $0.2*\cos(\al(y-y_{0}))$\\
\linebreak
So the curl of the wind forcing is given by a sine wave.\\
\linebreak
And thermal forcing:\\
There is currently no thermal forcing, just a pre-determined initial temperature.\\
\linebreak
The discretisation in space is given by:\\
\linebreak
And in time:\\
\linebreak
The ALE scheme is implemented according to:\\
Not currently available as output\\
\linebreak
The equation of state is:\\
linear; but density can be chosen as output in io\textunderscore meandata. \\
\linebreak
Boundary conditions implemented currently are:\\
No slip. Possible to enable free slip though (see namelist.oce). There is also bottom drag with amplitude defined in namelist.oce. This file also has parameterisations: currently harmonic leith with biharmonic background. the file oce\textunderscore dyn.f90 has further options for parameterisations.\\
\linebreak
Other possible boundary conditions:
\begin{itemize}
	\item No slip?\\
	\item Free slip?\\
	\item Partial slip?--Not available
\end{itemize}
Other parameters:
\begin{itemize}
	\item Viscosity.
	\item 
\end{itemize}
There is a forced rotation option in the namelist.config file. This has the effect of...\\
\linebreak
Is salinity disabled?\\
\linebreak
Are initial conditions from rest?\\
\linebreak
Are outputs monthly averages?\\
Yes\\
\linebreak
What is abg?\\
Rotation of mesh. Should be set to zero. \\
\linebreak
How do we get corresponding physical position from mesh data?


\end{document}
